---
title: "Statistical Inference Course Project"
author: "Daren Thomas"
date: "17 September 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 1: Simulation Excercise

In this project, I investigate the exponential distribution in R and compare it with the Central Limit Theorem. Let's start with a sample of the exponential distribution:

```{r}
n <- 1000 * 40  # taking n as the number of simulations of 40 averages...
lambda <- 0.2
sim <- rexp(n, lambda)
plot(sim)
```

What a mess! Let's clean that up a bit... what would, say, the distribution of averages of 40 exponentials look like?

```{r}
bin.size <- 40
sim.means <- apply(replicate(n / bin.size, sample(sim, size = bin.size)), 2, mean)
plot(sim.means)
```

Nice! We could also look at the distribution...

```{r}
hist(sim.means, breaks = 20, main = 'Distribution of means of 40 exponentials')
```

Let's calculate some simple statistics... say...

### Sample mean and theoretical mean

```{r}
hist(sim.means, breaks = 20, main = 'Distribution of means of 40 exponentials')
sample.mean <- mean(sim.means)
theoretical.mean <- 1/lambda

abline(v=sample.mean, col='blue')
abline(v=theoretical.mean, col='red')
legend("topright", c("Sample mean", "Theoretical mean"), fill=c("blue", "red"))
```

The sample mean is `r sample.mean` and the theoretical mean is `r theoretical.mean`. The difference is `r theoretical.mean - sample.mean`
