---
title: "Statistical Inference Course Project"
author: "Daren Thomas"
date: "17 September 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 1: Simulation Excercise

In this project, I investigate the exponential distribution in R and compare it with the Central Limit Theorem. Let's start with a sample of the exponential distribution:

```{r}
n <- 1000 * 40  # taking n as the number of simulations of 40 averages...
lambda <- 0.2
sim <- rexp(n, lambda)
plot(sim)
```

What a mess! Let's clean that up a bit... what would, say, the distribution of averages of 40 exponentials look like?

```{r}
bin.size <- 40
sim.means <- apply(replicate(n / bin.size, sample(sim, size = bin.size)), 2, mean)
plot(sim.means)
```

Nice! We could also look at the distribution...

```{r}
hist(sim.means, breaks = 20, main = 'Distribution of means of 40 exponentials')
```

Let's calculate some simple statistics... say...

### Sample mean and theoretical mean

```{r}
hist(sim.means, breaks = 20, main = 'Distribution of means of 40 exponentials')
sample.mean <- mean(sim.means)
theoretical.mean <- 1/lambda

abline(v=sample.mean, col='blue')
abline(v=theoretical.mean, col='red')
legend("topright", c("Sample mean", "Theoretical mean"), fill=c("blue", "red"))
```

The sample mean is `r sample.mean` and the theoretical mean is `r theoretical.mean`. The difference is `r theoretical.mean - sample.mean`. I use the mean of the exponential distribution as the mean of the distribution of the means because CLM states I can.

### Variance of sample vs. theoretical variance

The CLM states that the averages are approximately normal, with distributions centered at the population mean and with standard deviation equal to the standard error of the mean.

The Variance of the sample is `r var(sim.means)` while the theoretical variance is `r 1/lambda`. The difference is `r 1/lambda - var(sim.means)`.
